---
title: "RNAseq"
author: "Dani Blumsin"
da: "7/8/2022"
output: html_document
---

knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)

if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

devtools::install_github('kevinblighe/EnhancedVolcano')
BiocManager::install("DESeq2")
BiocManager::install("apeglm")
BiocManager::install("topGO")
BiocManager::install("BiocParallel")
library("remotes")
remotes::install_github("federicomarini/GeneTonic", 
                        dependencies = TRUE, build_vignettes = TRUE)
                        
```{r}
library("GeneTonic")
library("macrophage")
library("org.Hs.eg.db")
library("AnnotationDbi")
library(DESeq2)
library(dplyr)
library(foreach)
library(data.table)
library(splines)
library(ggthemes)
library(scales)
library(gridExtra)
library(tidyr)
library(pheatmap)
library(RColorBrewer)
library(ggplot2)
library(apeglm)
library(tidyverse)
library(topGO)
library(GO.db)
library(readr)
library(readxl)
library(lubridate)
library(patchwork)
library(EnhancedVolcano)
library(biomaRt)

```

```{r}
setwd("/Users/danielleblumstein/Documents/UNH/rnaseq/diet_final")
#call adams fancy functions
source("adams_fancy_functions.R")
ensembl = useMart( "ensembl", dataset = "hsapiens_gene_ensembl" )

dir.create("figures")
dir.create("figures/TalkGeneFigures")
dir.create("figures4publication")
dir.create("results")
dir.create("results/DEresults")
dir.create("results/GOresults")
dir.create("results/WGCNAresults")
```

```{r}
mappingdata_edit <- read.delim("~/Documents/UNH/rnaseq/mappingdata_edit.txt")

mappingdata_editM <- subset(mappingdata_edit, sex == 'M')
hist(mappingdata_editM$number.of.reads.mapped)

mappingdata_editF <- subset(mappingdata_edit, sex == 'F')
hist(mappingdata_editF$number.of.reads.mapped)

t.test(mappingdata_editM$number.of.reads.mapped,mappingdata_editF$number.of.reads.mapped)

mappingdata_editLF <- subset(mappingdata_edit, diet == 'lowfat')
hist(mappingdata_editLF$number.of.reads.mapped)

mappingdata_editSD <- subset(mappingdata_edit, diet == 'standard')
hist(mappingdata_editSD$number.of.reads.mapped)
```


make sample name spreadsheet
```{r}
### List all samples from expression data 
# get the directory/path for each sample in this study
base_dir <- getwd()

# get directories
filename <- list.files(path = "data/", pattern = "gene.counts", full.names = FALSE, recursive = FALSE)

# append full file path to each sample id
files <- file.path(base_dir, "data/", filename) # files = directory + salmon directory + sample name + quantifictaion file name
names(files) <- "data/"
all(file.exists(files)) # do these all actually exist?

# Make sample spreadsheet 
sample <- as.data.frame(filename)

# get sample
sample$individal <- filename  %>% gsub(pattern = "gene.counts", replacement = "")

###############come back to##############
# get tissue
sample$tissue <- unlist(lapply(strsplit(sample$individal, split = "_"),"[[", 3))

# get sex
sample$sex <- unlist(lapply(strsplit(sample$individal, split = "_"),"[[", 2))

# get trt
sample$trt <- unlist(lapply(strsplit(sample$individal, split = "_"),"[[", 4))

# get sample ID
sample$Animal_ID <- unlist(lapply(strsplit(sample$individal, split = "_"),"[[", 1))

#grabbing the phys data and electro/ weight data. currently comming from diff script but it happens to be in my global enviro. will need to change this 

cages <- read.csv("~/Documents/UNH/metabolic chamber/diet/cages") 
electrolyte_data <- read_excel("data/electrolyte_data.xlsx", na = "NA")
col_order <- c("Animal_ID", "StartTime", "VO2", "VCO2", "H2Omg","EE_kJH", "RQ", "weight", "seconds")
phys_data <- cages[, col_order]
electrolyte_data$Animal_ID <- as.character(electrolyte_data$mouse_ID)

ggplot(cages, aes(y = VO2, x = seconds, color = experiment))+
  geom_point()+
  geom_smooth()+
  facet_grid(~Sex)

##splitting out only the phys data between 11:00 and 12:00
noon <- period_to_seconds(hms("11:00:00")):period_to_seconds(hms("12:00:00"))
phys_data$TimeOfDay <- ifelse(phys_data$seconds %in% noon, "noon", "other")
phys_data <- phys_data[phys_data$TimeOfDay == "noon", ]

mean_VO2<- setNames(aggregate(phys_data$VO2, list(phys_data$Animal_ID), FUN=mean),c("Animal_ID", "mean_VO2"))
mean_VCO2<- setNames(aggregate(phys_data$VCO2, list(phys_data$Animal_ID), FUN=mean),c("Animal_ID", "mean_VCO2"))
mean_EE<- setNames(aggregate(phys_data$EE_kJH, list(phys_data$Animal_ID), FUN=mean),c("Animal_ID", "mean_EE"))
mean_RQ<- setNames(aggregate(phys_data$RQ, list(phys_data$Animal_ID), FUN=mean),c("Animal_ID", "mean_RQ"))
mean_H2Omg<- setNames(aggregate(phys_data$H2Omg, list(phys_data$Animal_ID), FUN=mean),c("Animal_ID", "mean_H2Omg"))

##merge all data frames together
samples_phys<- Reduce(function(x, y) merge(x, y, all=TRUE), list(mean_VO2, mean_VCO2, mean_RQ, mean_H2Omg, mean_EE,sample))
samples_phys$Animal_ID <- as.character(samples_phys$Animal_ID)
samples <- electrolyte_data %>% right_join(samples_phys)

samples$...6 <- NULL
samples$`experiment ` <- NULL
samples$experiment_date <- NULL
samples$cage_ID <- NULL
samples$cage_number<- NULL
samples$weight2<- NULL
samples$weight3<- NULL
samples$end_weight<- NULL
samples$diet<- NULL
samples$mouse_ID<- NULL

samples<- samples[!is.na(samples$filename),]

write.csv(samples, "results/samplespreadsheet.csv", row.names = FALSE)
#write.csv(cages, "results/cages.csv", row.names = FALSE)
```


```{r}
## Import annotation documents 
#annos <- read_delim("data/annos.txt", delim = "\t", 
#+     escape_double = FALSE, col_names = FALSE, 
#+     locale = locale(), trim_ws = TRUE)
annos <- read.csv("data/annos.txt")

library("org.Hs.eg.db")
annos$ensid = mapIds(org.Hs.eg.db,
                    keys=annos$gene, 
                    column="ENSEMBL",
                    keytype="SYMBOL",
                    multiVals="first")

#######################come back to ##########################
# make spreadsheet for deseq
#filenames <- samples[,21]
```


```{r}
countdata <- data.frame(annos$transcript)
colnames(countdata) <- "transcript"
for (counts in (1:length(samples$filename)))
  {
  # import data for that sample
  toimport <- samples$filename[counts]
  tmpdata <- read.table(paste0("data/", toimport), header = FALSE, sep = "\t")
  # sample id
  samplename <- toimport  %>% gsub(pattern = "gene.counts", replacement = "")
  samplename <- gsub("_S.*", "", samplename)
  colnames(tmpdata) <- c("transcript", samplename)
  countdata <- dplyr::left_join(countdata, tmpdata, by = "transcript")
}
dim(countdata)
countdata <- data.frame(countdata, row.names = 1)
names(countdata)=str_sub(names(countdata),2)

write.csv(countdata, "countdata.csv", row.names = FALSE)
```


Quick data control. Change any "NA" values to "0"
```{r data QC}
# change all NA to 0
print(paste0("Number of NA after import:", sum(is.na(countdata))))
countdata[is.na(countdata)] <-  0
print(paste0("Number of NA after NA removal:", sum(is.na(countdata))))
print(paste0("Number of genes after import:", nrow(countdata)))

## remove all unknown proteins
countdata <- countdata[!rownames(countdata) == "proteins of unknown function", ]
print(paste("Number of genes after removing all 'proteins of unknown function':", nrow(countdata)))

dim(countdata)
```


The count data table is currently a dataframe of counts, with each row representing a transcript in the genome. Some genes are represented by multiple transcripts in the genome, and hence our count data table. Here I will manipulate the data so that we can do a gene level analysis.

```{R Create a data frame of gene level count data}
# make rownames column 1
countdata <- setDT(countdata, keep.rownames = "transcript")[]

# add gene symbol
countdata2 <- dplyr::left_join(annos, countdata, by = "transcript")
dim(countdata2)

# change all NA to 0
print(paste("Number of NA after import:", sum(is.na(countdata2))))

countdata2 <- countdata2 %>% 
    mutate_at(5:193, ~replace_na(.,0))

# combine all counts that map to the same gene
countdata3 <- aggregate(countdata2[, 5:193], list(countdata2$gene), sum)
dim(countdata3)

countdata3[is.na(countdata3)] <-  0
colnames(countdata3)[1] <- "Gene"
countdata <- countdata3

# make gene name column
countdata <- data.frame(countdata, row.names = 1)
print(paste("Number of genes after combining all counts that map to the same gene:", nrow(countdata)))

for ( col in 1:ncol(countdata)){
    colnames(countdata)[col] <-  sub("X*", "", colnames(countdata)[col])
}

# save gene-level count data
write.table(countdata, "data/gene.level.count.data.tsv", row.names = TRUE, sep = "\t")
```


t tests
```{r}
weight_F_lowfat <-samples %>% filter(sex == "F") 
weight_F_lowfat <-weight_F_lowfat %>% filter(trt == "lowfat") 
weight_F_lowfat = weight_F_lowfat[!duplicated(weight_F_lowfat$Animal_ID),]
weight_F_standard <-samples %>% filter(sex == "F") 
weight_F_standard <-weight_F_standard %>% filter(trt == "standard") 
weight_F_standard = weight_F_standard[!duplicated(weight_F_standard$Animal_ID),]
t.test(weight_F_standard$weight, weight_F_lowfat$weight)

weight_M_lowfat <-samples %>% filter(sex == "M") 
weight_M_lowfat <-weight_M_lowfat %>% filter(trt == "lowfat") 
weight_M_lowfat = weight_M_lowfat[!duplicated(weight_M_lowfat$Animal_ID),]
weight_M_standard <-samples %>% filter(sex == "M") 
weight_M_standard <-weight_M_standard %>% filter(trt == "standard") 
weight_M_standard = weight_M_standard[!duplicated(weight_M_standard$Animal_ID),]
t.test(weight_M_standard$weight, weight_M_lowfat$weight)

weight_lowfat <-samples %>% filter(trt == "lowfat") 
weight_lowfat = weight_lowfat[!duplicated(weight_lowfat$Animal_ID),]
weight_standard <-samples %>% filter(trt == "standard")
weight_standard = weight_standard[!duplicated(weight_standard$Animal_ID),]
t.test(weight_lowfat$weight, weight_standard$weight)

weight_F <-samples %>% filter(sex == "F") 
weight_F = weight_F[!duplicated(weight_F$Animal_ID),]
weight_M <-samples %>% filter(sex == "M")
weight_M = weight_M[!duplicated(weight_M$Animal_ID),]
t.test(weight_F$weight, weight_M$weight)

meanEE_F_lowfat <-samples %>% filter(sex == "F") 
meanEE_F_lowfat <-meanEE_F_lowfat %>% filter(trt == "lowfat") 
meanEE_F_lowfat = meanEE_F_lowfat[!duplicated(meanEE_F_lowfat$Animal_ID),]
meanEE_F_standard <-samples %>% filter(sex == "F") 
meanEE_F_standard <-meanEE_F_standard %>% filter(trt == "standard") 
meanEE_F_standard = meanEE_F_standard[!duplicated(meanEE_F_standard$Animal_ID),]
t.test(meanEE_F_standard$mean_EE, meanEE_F_lowfat$mean_EE)

meanEE_lowfat <-samples %>% filter(trt == "lowfat") 
meanEE_lowfat = meanEE_lowfat[!duplicated(meanEE_lowfat$Animal_ID),]
meanEE_standard <-samples %>% filter(trt == "standard") 
meanEE_standard = meanEE_standard[!duplicated(meanEE_standard$Animal_ID),]
t.test(meanEE_lowfat$mean_EE, meanEE_standard$mean_EE)

meanEE_F <-samples %>% filter(sex == "F")
meanEE_F = meanEE_F[!duplicated(meanEE_F$Animal_ID),]
meanEE_M <-samples %>% filter(sex == "M")
meanEE_M = meanEE_M[!duplicated(meanEE_M$Animal_ID),]
t.test(meanEE_F$mean_EE, meanEE_M$mean_EE) 

meanEE_M_lowfat <-samples %>% filter(sex == "M") 
meanEE_M_lowfat <-meanEE_M_lowfat %>% filter(trt == "lowfat") 
meanEE_M_lowfat = meanEE_M_lowfat[!duplicated(meanEE_M_lowfat$Animal_ID),]
meanEE_M_standard <-samples %>% filter(sex == "M") 
meanEE_M_standard <-meanEE_M_standard %>% filter(trt == "standard") 
meanEE_M_standard = meanEE_M_standard[!duplicated(meanEE_M_standard$Animal_ID),]
t.test(meanEE_M_standard$mean_EE, meanEE_M_lowfat$mean_EE)

meanRQ_F_lowfat <-samples %>% filter(sex == "F") 
meanRQ_F_lowfat <-meanRQ_F_lowfat %>% filter(trt == "lowfat") 
meanRQ_F_lowfat = meanRQ_F_lowfat[!duplicated(meanRQ_F_lowfat$Animal_ID),]
meanRQ_F_standard <-samples %>% filter(sex == "F") 
meanRQ_F_standard = meanRQ_F_standard[!duplicated(meanRQ_F_standard$Animal_ID),]
meanRQ_F_standard <-meanRQ_F_standard %>% filter(trt == "standard") 
t.test(meanRQ_F_standard$mean_RQ, meanRQ_F_lowfat$mean_RQ)

meanRQ_M_lowfat <-samples %>% filter(sex == "M") 
meanRQ_M_lowfat <-meanRQ_M_lowfat %>% filter(trt == "lowfat") 
meanRQ_M_lowfat = meanRQ_M_lowfat[!duplicated(meanRQ_M_lowfat$Animal_ID),]
meanRQ_M_standard <-samples %>% filter(sex == "M") 
meanRQ_M_standard <-meanRQ_M_standard %>% filter(trt == "standard") 
meanRQ_M_standard = meanRQ_M_standard[!duplicated(meanRQ_M_standard$Animal_ID),]
t.test(meanRQ_M_standard$mean_RQ, meanRQ_M_lowfat$mean_RQ)

meanRQ_lowfat <-samples %>% filter(trt == "lowfat") 
meanRQ_lowfat = meanRQ_lowfat[!duplicated(meanRQ_lowfat$Animal_ID),]
meanRQ_standard <-samples %>% filter(trt == "standard") 
meanRQ_standard = meanRQ_standard[!duplicated(meanRQ_standard$Animal_ID),]
t.test(meanRQ_lowfat$mean_RQ, meanRQ_standard$mean_RQ)

meanRQ_F <-samples %>% filter(sex == "F") 
meanRQ_F = meanRQ_F[!duplicated(meanRQ_F$Animal_ID),]
meanRQ_M <-samples %>% filter(sex == "M")
meanRQ_M = meanRQ_M[!duplicated(meanRQ_M$Animal_ID),]
t.test(meanRQ_F$mean_RQ, meanRQ_M$mean_RQ)

meanH2O_F_lowfat <-samples %>% filter(sex == "F") 
meanH2O_F_lowfat <-meanH2O_F_lowfat %>% filter(trt == "lowfat") 
meanH2O_F_lowfat = meanH2O_F_lowfat[!duplicated(meanH2O_F_lowfat$Animal_ID),]
meanH2O_F_standard <-samples %>% filter(sex == "F") 
meanH2O_F_standard <-meanH2O_F_standard %>% filter(trt == "standard") 
meanH2O_F_standard = meanH2O_F_standard[!duplicated(meanH2O_F_standard$Animal_ID),]
t.test(meanH2O_F_standard$mean_H2Omg, meanH2O_F_lowfat$mean_H2Omg)

meanH2O_M_lowfat <-samples %>% filter(sex == "M") 
meanH2O_M_lowfat <-meanH2O_M_lowfat %>% filter(trt == "lowfat") 
meanH2O_M_lowfat = meanH2O_M_lowfat[!duplicated(meanH2O_M_lowfat$Animal_ID),]
meanH2O_M_standard <-samples %>% filter(sex == "M") 
meanH2O_M_standard <-meanH2O_M_standard %>% filter(trt == "standard") 
meanH2O_M_standard = meanH2O_M_standard[!duplicated(meanH2O_M_standard$Animal_ID),]
t.test(meanH2O_M_standard$mean_H2Omg, meanH2O_M_lowfat$mean_H2Omg)

meanH2Omg_lowfat <-samples %>% filter(trt == "lowfat") 
meanH2Omg_lowfat = meanH2Omg_lowfat[!duplicated(meanH2Omg_lowfat$Animal_ID),]
meanH2Omg_standard <-samples %>% filter(trt == "standard") 
meanH2Omg_standard = meanH2Omg_standard[!duplicated(meanH2Omg_standard$Animal_ID),]
t.test(meanH2Omg_lowfat$mean_H2Omg, meanH2Omg_standard$mean_H2Omg)

meanH2Omg_F <-samples %>% filter(sex == "F") 
meanH2Omg_F = meanH2Omg_F[!duplicated(meanH2Omg_F$Animal_ID),]
meanH2Omg_M <-samples %>% filter(sex == "M")
meanH2Omg_M = meanH2Omg_M[!duplicated(meanH2Omg_M$Animal_ID),]
t.test(meanH2Omg_F$mean_H2Omg, meanH2Omg_M$mean_H2Omg)
```

### DESeq data creation 

```{r DESEQ creation}
dim(countdata)
dim(samples)
# crea DESeq dataset
countdata <- countdata+1
dds <- DESeqDataSetFromMatrix(countData = countdata,
                                       colData = samples,
                                       design = ~ sex + tissue + trt)


table(samples['sex'])
table(samples['tissue'])
table(samples['trt'])

sum(samples$trt== 'standard' & samples$tissue == "gi")
sum(samples$trt== 'standard' & samples$tissue == "liv")
sum(samples$trt== 'standard' & samples$tissue == "lu")
sum(samples$trt== 'standard' & samples$tissue == "kid")
sum(samples$trt== 'standard' & samples$tissue == "hyp")

sum(samples$trt== 'lowfat' & samples$tissue == "gi")
sum(samples$trt== 'lowfat' & samples$tissue == "liv")
sum(samples$trt== 'lowfat' & samples$tissue == "lu")
sum(samples$trt== 'lowfat' & samples$tissue == "kid")
sum(samples$trt== 'lowfat' & samples$tissue == "hyp")

sum(samples$sex== 'M' & samples$tissue == "gi")
sum(samples$sex== 'M' & samples$tissue == "liv")
sum(samples$sex== 'M' & samples$tissue == "lu")
sum(samples$sex== 'M' & samples$tissue == "kid")
sum(samples$sex== 'M' & samples$tissue == "hyp")

sum(samples$sex== 'F' & samples$tissue == "gi")
sum(samples$sex== 'F' & samples$tissue == "liv")
sum(samples$sex== 'F' & samples$tissue == "lu")
sum(samples$sex== 'F' & samples$tissue == "kid")
sum(samples$sex== 'F' & samples$tissue == "hyp")
```


### Crea a PCA

```{R PCA}
vsd <- varianceStabilizingTransformation(dds, blind = FALSE)
# get PC1 and PC2 data
pcaData <- plotPCA(vsd, intgroup = c("tissue","trt"), returnData = TRUE)
# get percent variation
percentVar <- round(100 * attr(pcaData, "percentVar"))
# pca code
pca <- ggplot(pcaData, aes(x = PC1, y = PC2, color = tissue, shape = trt, name=name)) +
  stat_ellipse(aes(group = tissue, linetype = tissue), type = "t", level = 0.95, size = 1.25, show.legend = FALSE) +
  #scale_linetype_manual(values=c("twodash", "longdash", "solid"), guide = FALSE) +
  geom_point(size = 3, show.legend = TRUE) + 
  #scale_color_manual(values = c("orange1", "red1", "yellow2", "chartreuse4")) +
  #scale_shape_manual(values = c(8,9,15,16,10,17,18,11,13)) + 
  xlab(paste0("PC1: ", percentVar[1], "% variance")) +
  ylab(paste0("PC2: ", percentVar[2], "% variance")) +
  coord_fixed() + theme_bw() +
  guides(shape = guide_legend(order = 1),color = guide_legend(order = 2)) 
# save
ggsave("figures/PCA.png", width = 8, height = 4.6, dpi = 600)


library(plotly)
ggplotly(pca)

p <- ggplotly(pca)
```


### Make the final PCA for the publication

```{R PCA for publication}
# make sure levels of morphs are correctly specified for colors
# pca code
ggplot(pcaData, aes(x = PC1, y = PC2, color = tissue)) +
  stat_ellipse(aes(group = tissue, linetype = tissue), type = "t", level = 0.95, size = 1.25, show.legend = FALSE) +
  #scale_linetype_manual(values=c("twodash", "longdash", "solid"), guide = FALSE) +
  geom_point(size = 3, show.legend = TRUE) + 
  #scale_color_manual(values = c("orange1", "red1", "yellow2", "chartreuse4")) +
  #scale_shape_manual(values = c(8,9,15,16,10,17,18,11,13)) + 
  xlab(paste0("PC1: ", percentVar[1], "% variance")) +
  ylab(paste0("PC2: ", percentVar[2], "% variance")) +
  coord_fixed() + theme_bw() +
  guides(shape = guide_legend(order = 1),color = guide_legend(order = 2)) 
  #annota("xt", label = "Ranitomeya fantastica", x = 45, y = -22.5, size = 4, colour = "black", fontface = "italic") +
  #annota("xt", label = "Ranitomeya imitator", x = 38, y = 22, size = 4, colour = "black", fontface = "italic") +
  #annota("xt", label = "Ranitomeya variabilis", x = -45, y = 12, size = 4, colour = "black", fontface = "italic") 
  
ggsave("figures4publication/PCA.png", width = 8, height = 4.6, dpi = 600)
```


### Crea a heatmap

Make a heatmap with rows/columns ordered by populations then time?
#doesn't do much for my data yet

```{r HEATWAVE}
# crea sample distances and sample distance matrix
sampleDists <- dist(t(assay(vsd)))
sampleDistMatrix <- as.matrix( sampleDists )
# rename columns/rows so they are inrpretable
#rownames(sampleDistMatrix) <- gsub(patrn = "_M.*$", replacement = "", rownames(sampleDistMatrix))
#colnames(sampleDistMatrix) <- gsub(patrn = "_M.*$", replacement = "", colnames(sampleDistMatrix))
# plot heatmap
heatmapcolors <- colorRampPalette( rev(brewer.pal(9, "Blues")) )(255)
pheatmap(sampleDistMatrix, clusr_rows = FALSE, clusr_cols = FALSE, col = heatmapcolors)
# plot and save heatmap
png("figures/heatmap.png", width = 14.63, height = 10, units = "in", res = 600)
pheatmap(sampleDistMatrix, clusr_rows = FALSE, clusr_cols = FALSE, col = heatmapcolors)
```

### Differential expression analyses

I will run some differential expression analyses now. I want to account for sex, trt, and tissue in my analyses.


Run differential expression analysis. What is the effect of sex?

```{r LRT sex}
ddssex <- DESeqDataSetFromMatrix(countData = countdata,
                                       colData = samples,
                                       design = ~ trt + tissue + sex)
# run LRT test
ddssex <- DESeq(ddssex, test="LRT", reduced = ~ trt + tissue)
ressex <- results(ddssex)
ressex$transcript <- mcols(ressex)$transcript
# how many are "significant"?
table(ressex[,"padj"] < 0.05)
```

```{r LRT tissue}
ddstissue <- DESeqDataSetFromMatrix(countData = countdata,
                                       colData = samples,
                                       design = ~ trt + sex + tissue)
# run LRT test
ddstissue <- DESeq(ddstissue, test="LRT", reduced = ~ trt + sex)
restissue <- results(ddstissue)
restissue$symbol <- mcols(restissue)$symbol
# how many are "significant"?
table(restissue[,"padj"] < 0.05)
```

```{r LRT trt}
ddstrt <- DESeqDataSetFromMatrix(countData = countdata,
                                       colData = samples,
                                       design = ~ sex + tissue + trt)

ddstrt <- estimateSizeFactors(ddstrt)
idx <- rowSums(counts(ddstrt, normalized=TRUE) >= 10 ) >= 8

ddstrt <- ddstrt[idx,]
print(paste("Number of genes after filtering by total expression:", nrow(ddstrt)))

# run LRT test
ddstrt <- DESeq(ddstrt, test="LRT", reduced = ~ sex + tissue)
restrt <- results(ddstrt)
restrt$symbol <- mcols(restrt)$symbol
# how many are "significant"?
table(restrt[,"padj"] < 0.05)

ddstrt <- DESeq(ddstrt,test="Wald")
annos$ensembl <- as.character(annos$ensid)

#rename transcript rows to ensid names
rownames(ddstrt)=annos$ensid[match(names(ddstrt),annos$gene)]
#remove the NAs (those transcripts dont have a ensid name)
ddstrt <- ddstrt[complete.cases(rownames(ddstrt)), ]
ddstrt <- ddstrt[!rownames(ddstrt) == "NULL", ] 
rownames(ddstrt)=annos$gene[match(names(ddstrt),annos$ensid)]

ddstrt <- estimateSizeFactors(ddstrt)
normalized_counts <- vst(ddstrt, blind = FALSE)
all_Expr <- assay(normalized_counts)
```


```{r}
require(GGally)
library(vegan)

cca_cols <- c("kid_lowfat"="#E3EB94","kid_standard" = "#D0E14E","lu_lowfat"="#94BEBD","lu_standard"="#48918B", "liv_lowfat"="#A9DAAF","liv_standard"="#6EBF75", "hyp_lowfat"="#A47FC7","hyp_standard"="#6327A2", "gi_lowfat" = "#959BB8","gi_standard"="#414D86", "in bound" = "lightgrey" ,"outlier"= "darkgrey")

#genes_for_cca <- read.table("~/Documents/UNH/rnaseq/diet_final/genes_for_cca.csv", quote="\"", comment.char="")
genes_for_cca <- rownames(countdata)
all_Expr<- as.data.frame(all_Expr)

#subset_all_Expr <- all_Expr %>% filter(rownames(all_Expr) %in% genes_for_cca$X1)

t_subset_all_Expr <-t(all_Expr)
samples<-as.data.frame(samples)
traitData_V2 <- samples %>% dplyr::select(c("trt","weight","mean_RQ","mean_EE","mean_H2Omg","individal","tissue", "sex","Animal_ID"))

#traitData_V2 <- traitData_V2 |>
 #   mutate(tt = fct_cross(trt, sex, tissue))

traitData_V2$trt <- as.factor(traitData_V2$trt)
traitData_V2$sex <- as.factor(traitData_V2$sex)
rownames(traitData_V2) <- traitData_V2$individal
traitData_V2 <- subset(traitData_V2, select = -individal)
traitData_V2$tissue <- as.factor(traitData_V2$tissue)
traitData_V2$Animal_ID <- as.factor(traitData_V2$Animal_ID)

#ggpairs(traitData_V2)

cca_model <- cca(t_subset_all_Expr ~ trt+weight+mean_RQ+mean_EE+mean_H2Omg +sex +tissue,data = traitData_V2)

#anova the model to find the significance of the terms
#anova(cca_model)
#anova(cca_model,by="terms")
#anova(cca_model,by="axis")
#anova(cca_model, by = "margin")

#Get CCA scores
df_species  <- data.frame(summary(cca_model)$species[,1:2])# get the genes CC1 and CC2 scores

sd <- sd(df_species$CCA1)
out_ind1 <- df_species %>% filter(CCA1 > sd*2)
out_ind2 <- df_species %>% filter(CCA1 < -(sd*2))

sd <- sd(df_species$CCA2)
out_ind3 <- df_species %>% filter(CCA2 > sd*3)
out_ind4 <- df_species %>% filter(CCA2 < -(sd*3))

#put the 3sd (99% confidence) from the genes are in a GO analysis
out <- distinct(rbind(out_ind1,out_ind2,out_ind3,out_ind4))

write.table(out, "results/cca_outliers.csv", row.names = TRUE, sep = "\t")
out$color <- "outlier"

norm <- df_species[!(row.names(df_species) %in% rownames(out)),]
norm$color <- "in bound"

graphing_data <- as.data.frame(rbind(norm,out))

df_sites  <- data.frame(summary(cca_model)$sites[,1:2])# get the indivisuals CC1 and CC2 scores
df_environ  <- data.frame(scores(summary(cca_model)$biplot[c(2,3,4),1:2])) #get the physio vars CC1 and CC2 scores

df_tissue  <- scores(summary(cca_model)$centroids[5:9,1:2]) #get the tissue centriod vars CC1 and CC2 scores


rc <- function (u) paste0(u[c(3,4)], collapse = "_")
#make col names just the tissue, sex, trt
df_sites$lable <- unlist(lapply(strsplit(rownames(df_sites), split = "_"),rc))

CCA1_varex<-round(summary(cca_model)$cont$importance[2,1]*100,2) #Get percentage of variance explained by first axis
CCA2_varex<-round(summary(cca_model)$cont$importance[2,2]*100,2) #Get percentage of variance explained by second axis

#Set a scaling variable to multiply the CCA values, in order to get a very similar plot to the the one generated by plot(cca_model). You can adjust it according to your data
scaling_factor <- 2
library(paletteer)

cca_plot1 <- ggplot(df_species, 
       aes(x=CCA1, y=CCA2)) + 
  #Draw lines on x = 0 and y = 0
  geom_hline(yintercept=0, 
             linetype="dashed") +
  geom_vline(xintercept=0, 
             linetype="dashed") +
  #coord_fixed()+
  #Add species text
    geom_point(data=graphing_data, 
            aes(x=CCA1,#Score in CCA1 to add species text
                y=CCA2,#Score in CCA2 to add species text
                hjust=0.5*(1-sign(CCA1)),#Set the text horizontal alignment according to its position in the CCA plot
                vjust=0.5*(1-sign(CCA2)),
                color=factor(color)),#Set the text vertical alignment according to its position in the CCA plot
            #color = c("grey","#160b39"),
            size = 1)+
  #Add sites text
  geom_point(data=df_sites, 
            aes(col = df_sites$lable,
                x=CCA1,#Score in CCA1 to add species text
                y=CCA2))+#Score in CCA2 to add species text
  #Add environmental vars arrows
  geom_segment(data=df_environ, 
               aes(x=0, #Starting coordinate in CCA1 = 0 
                   xend=CCA1*scaling_factor,#Ending coordinate in CCA1  
                   y=0, #Start in CCA2 = 0
                   yend=CCA2*scaling_factor), #Ending coordinate in CCA2 
               color="black", #set color
               arrow=arrow(length=unit(0.01,"npc"))#Set the size of the lines that form the tip of the arrow
               )+
  #Add environmental vars text
  geom_text(data=df_environ, 
            aes(x=CCA1*scaling_factor, 
                y=CCA2*scaling_factor,
                label=rownames(df_environ),
                hjust=0.5*(1-sign(CCA1)),#Add the text of each environmental var at the end of the arrow
                vjust=0.5*(1-sign(CCA2))),#Add the text of each environmental var at the end of the arrow 
            color="black")+
  #Set bw theme
  theme+
  scale_color_manual(values=cca_cols, name = df_sites$lable)+
  #xlim(-0.8, 0.7)+
  #ylim(-0.5, 0.7)+
  #Set x and y axis titles
  labs(x=paste0("CCA1 (",CCA1_varex," %)"),
       y=paste0("CCA2 (",CCA2_varex," %)"))+
  theme(legend.title=element_blank())
  #theme(legend.position = "none")

cca_plot1

ggsave("cca_plot.png", width = 7, height = 7, dpi = 600)
```


```{r}
go_kid <- go_kid %>% rename(GO = V1,description = V2,kid_up = V3.x,kid_down=V3.y)
go_kid <- go_kid %>% rename(kid_up_pval = V4.x, kid_down_pval = V4.y)
go_kid <-select(go_kid, -tissue.x)
go_kid <-select(go_kid, -tissue.y)
go_kid <-select(go_kid, -DE.x)
go_kid <-select(go_kid, -DE.y)

go_gi <- go_gi %>% rename(GO = V1,description = V2,gi_up = V3.x,gi_down=V3.y)
go_gi <- go_gi %>% rename(gi_up_pval = V4.x, gi_down_pval = V4.y)
go_gi <-select(go_gi, -tissue.x)
go_gi <-select(go_gi, -tissue.y)
go_gi <-select(go_gi, -DE.x)
go_gi <-select(go_gi, -DE.y)

go_hyp <- go_hyp %>% rename(GO = V1,description = V2,hyp_up = V3.x,hyp_down=V3.y)
go_hyp <- go_hyp %>% rename(hyp_up_pval = V4.x, hyp_down_pval = V4.y)
go_hyp <-select(go_hyp, -tissue.x)
go_hyp <-select(go_hyp, -tissue.y)
go_hyp <-select(go_hyp, -DE.x)
go_hyp <-select(go_hyp, -DE.y)

go_liv <- go_liv %>% rename(GO = V1,description = V2,liv_up = V3.x,liv_down=V3.y)
go_liv <- go_liv %>% rename(liv_up_pval = V4.x, liv_down_pval = V4.y)
go_liv <-select(go_liv, -tissue.x)
go_liv <-select(go_liv, -tissue.y)
go_liv <-select(go_liv, -DE.x)
go_liv <-select(go_liv, -DE.y)

go_lu <- go_lu %>% rename(GO = V1,description = V2,lu_up = V3.x,lu_down=V3.y)
go_lu <- go_lu %>% rename(lu_up_pval = V4.x, lu_down_pval = V4.y)
go_lu <-select(go_lu, -tissue.x)
go_lu <-select(go_lu, -tissue.y)
go_lu <-select(go_lu, -DE.x)
go_lu <-select(go_lu, -DE.y)


go_list <- list(go_kid,go_gi,go_hyp,go_liv,go_lu)
go <- go_list %>% reduce(full_join, by=c("GO","description"))
go <- reshape2::melt(go,measure.vars = c("kid_up", "kid_down","gi_down","hyp_up","hyp_down","liv_up", "liv_down", "lu_up", "lu_down"))
go <- go %>% rename(count = value, tissue = variable)
go <- reshape2::melt(go,measure.vars = c("kid_up_pval","kid_down_pval","gi_down_pval","hyp_up_pval","hyp_down_pval","liv_up_pval","liv_down_pval","lu_up_pval","lu_down_pval"))
go <- go %>% rename(pval = value)
go<- within(go, tissue<-data.frame(do.call('rbind', strsplit(as.character(tissue), '_', fixed=TRUE))))
go$pval <- as.numeric(go$pval)
#go$pval <- round(go$pval ,digit=5)
go$count <- as.numeric(go$count)
go_clean <- na.omit(go)
up_go <- na.omit(filter(go, tissue$X2=="up"))
down_go <- na.omit(filter(go, tissue$X2=="down"))

theme<-theme(panel.background = element_blank(),panel.border=element_blank(),axis.line = element_line(colour = "black"),panel.grid.major = element_blank(),panel.grid.minor = element_blank(),
             strip.background=element_blank(),axis.text.x=element_text(colour="black",size=10),axis.text.y=element_text(colour="black",size=10),
             axis.ticks=element_line(colour="black"),plot.margin=unit(c(1,1,1,1),"line"))

up <-ggplot(up_go, aes(x = tissue$X1, y = description, color = tissue$X1, size = count)) +
  geom_point()+
  theme+
  ggtitle("up regulated")+
  theme(legend.position = "none")

down <- ggplot(down_go, aes(x = tissue$X1, y = description, color = tissue$X1, size = count)) +
  geom_point()+
  theme+
  ggtitle("down regulated")+
  theme(legend.position = "none")

up|down

ggsave("DE_GO_sex.png", width = 15, height = 10, dpi = 600)
```



```{r}
big_data_kid <- dplyr::filter(big_data_kid, grepl('GO', V1))
big_data_kid$tissue <- "kid"
rownames(big_data_kid) <- 1:nrow(big_data_kid)
left_join(big_data_kid, net_res, by = c("V2" = "count"))
go_wgcna_clean_kid_subset<-big_data_kid[c(1,2,3,4,5,6,8,9,13,17,18,20,23,27,31,34,35,37,38,40,41,43,46,47,49,51,52,53,54,55,57,58,59,60,61,62,62,67,68,72,73,74,77,80,81,82,85,86,87,88,89,91,92,93,95,96,97,98,108,109,110,111,112,113,114,115,116,119,120),]

big_data_liv <- dplyr::filter(big_data_liv, grepl('GO', V1))
big_data_liv$tissue <- "liv"
rownames(big_data_liv) <- 1:nrow(big_data_liv)
left_join(big_data_liv, net_res, by = c("V2" = "count"))
go_wgcna_clean_liv_subset<-big_data_liv[c(3,4,5,9,10,11,14,15,16,17,18,19,21,28,29,32,33,34,35,36,37,38,39,40,41,42,43,46,50,51,55,56,57,58,59,60,61,62,63,64,65,66,67,68,69,70,71,72,73,75,77,79,80,81,82,83,84,85,86,87),]

big_data_gi <- dplyr::filter(big_data_gi, grepl('GO', V1))
big_data_gi$tissue <- "gi"
rownames(big_data_gi) <- 1:nrow(big_data_gi)
left_join(big_data_gi, net_res, by = c("V2" = "count"))
go_wgcna_clean_gi_subset<-big_data_gi[c(1,2,3,5,6,7,9,11,12,13,14,15,19,21,22,23,25,27,28,32,33,34,36,37,38,39,40,43,46,48,49,50,51,52,54,55,56,58,59,60,61,62,63,64,65,70,75,76,78,79,80,81,82,83,96,99,107,111,112,113,114,115,117,119,120,121,122,123,124),]

big_data_hyp <- dplyr::filter(big_data_hyp, grepl('GO', V1))
big_data_hyp$tissue <- "hyp"
rownames(big_data_hyp) <- 1:nrow(big_data_hyp)
left_join(big_data_hyp, net_res, by = c("V2" = "count"))
go_wgcna_clean_hyp_subset<-big_data_hyp[c(1,6,7,8,10,13,14,15,16,17,21,23,24,25,37,45,48,49,50,52,53,56,55,57,59,60, 61,64,66,67,69,76,77,79,83,86,87,89,90,91,92,93,94,95,96,97,98,100,101),]

big_data_lu <- dplyr::filter(big_data_lu, grepl('GO', V1))
big_data_lu$tissue <- "lu"
rownames(big_data_lu) <- 1:nrow(big_data_lu)
left_join(big_data_lu, net_res, by = c("V2" = "count"))
go_wgcna_clean_lu_subset<-big_data_lu[c(1,2,6,8,9,10,13,14,17,18,19,20,21,23,25,26,29,30,32,34,35,36,38,44,45,47,48,49,53,54,55,57,58,60,61,62,63,64,66,67,68,69,73,81,82,90,91,92,93,94, 96,97,98,99,100,101,102,103,104,105,106,107,108,109,110,111,112,113,112,125,126,127,128,129,131,133,134,137,140,142,143,144,154,146,147,150,151,152,153),]

wgcna_data <- rbind(go_wgcna_clean_lu_subset,go_wgcna_clean_hyp_subset,go_wgcna_clean_gi_subset,go_wgcna_clean_liv_subset,go_wgcna_clean_kid_subset)
 
#left_join(big_data_lu, net_res, by = c("V2" = "count"))
coul = c(
              "trt" = "#fdc328",
              "iCa"="#f79342",
              "weight"="#e4695e",
              "Crea"="#c23c81",
              "mean_RQ"="#920fa3", 
              "AnGap"="#6c00a8",
              "mean_H2Omg"="#1fa088", 
              "Cl" = "#31688e",
              "BUN"="#886EA0", 
              "K" = "darkgrey" ,
              "TCO2"= "#70809c",
              "Hct"= "plum",
          "mean_EE"="#4cc26c",
         "sex"="purple")

cols <- c("kid"="#E3EB94","liv"="#A9DAAF","lu"="#94BEBD", "hyp"="#A47FC7", "gi" = "#959BB8")

#library(igraph)
links <-data.frame(
    source=wgcna_data$V2,
    target=wgcna_data$tissue,
    importance=wgcna_data$V3,
    edge = wgcna_data$pheno
    )

sort(table(c(links$source, links$target)), decreasing = TRUE)
result <- rle(sort(c(links[,1], links[,2])))
net_res <- data.frame(dot=result$lengths, count=result$values)

nodes <- dplyr::select(wgcna_data, V2,tissue)
nodes <- nodes %>%
  arrange(V2) %>%
  filter(duplicated(V2) == FALSE)


nodes[nrow(nodes) + 1,] = c("lu", "tissue")
nodes[nrow(nodes) + 1,] = c("liv","tissue")
nodes[nrow(nodes) + 1,] = c("gi","tissue")
nodes[nrow(nodes) + 1,] = c("hyp","tissue")
nodes[nrow(nodes) + 1,] = c("kid","tissue")

my_edge <- na.omit(as.numeric(wgcna_data$V3))
network <- graph_from_data_frame(d=links, vertices=nodes, directed=F) 


E(network)$color <- coul[as.numeric(as.factor(links$edge))]
E(network)$name <- as.factor(links$edge)
V(network)$color <- cols[as.numeric(as.factor(V(network)$tissue))]


## Function to wrap long strings
# Source: http://stackoverflow.com/a/7367534/496488
wrap_strings <- function(vector_of_strings,width){
  as.character(sapply(vector_of_strings, FUN=function(x){
                        paste(strwrap(x, width=width), collapse="\n")
                        }))
  }

# Apply the function to wrap the node labels
V(network)$name = wrap_strings(V(network)$name, 12)
network <- simplify(network, remove.multiple = F, remove.loops = T) 
network <- set_graph_attr(network, "layout", layout_with_kk(network))

library(qgraph)
e <- get.edgelist(network,names=FALSE)
l <- qgraph.layout.fruchtermanreingold(e,vcount=vcount(network),area=6*(vcount(network)^2),repulse.rad=(vcount(network)^3.1))

png("net.png", 4600, 3600,res= 1000)
par(mar=c(0,0,0,0))
plot(network,
    vertex.label.color="black",
    vertex.frame.color="transparent",
    vertex.size=log2(my_edge),
    vertex.label.cex = .2, 
    #vertex.label.dist = -1,
    edge.width=1,
    layout=l)
    
legend("bottomleft", legend=levels(as.factor(E(network)$name)), col = coul , bty = "n", pch=20 , pt.cex = .3, cex = .3, text.col="black" , horiz = FALSE, inset = c(0.1, 0.1))

dev.off()

```


```{r}
rownames(out)
library(gprofiler2)
#library(gProfileR)

gostres <- gost(query = rownames(out), 
                organism = "hsapiens", ordered_query = FALSE, 
                multi_query = FALSE, significant = TRUE, exclude_iea = FALSE, 
                measure_underrepresentation = FALSE, evcodes = FALSE, 
                user_threshold = 0.05, correction_method = "g_SCS", 
                domain_scope = "annotated", custom_bg = NULL, 
                numeric_ns = "", sources = NULL, as_short_link = FALSE, highlight = TRUE)

gostplot(gostres, capped = TRUE, interactive = TRUE)



```

res_diet
```{r}
#restrt_lu
#restrt_liv
#restrt_gi
#restrt_hyp
#restrt_kid
library(RColorBrewer)

summary(restrt_kid)
summary(restrt_hyp)
summary(restrt_gi)
summary(restrt_liv)
summary(restrt_lu)


kid_lfc <- as.data.frame(cbind(rownames(restrt_kid),restrt_kid$padj, restrt_kid$log2FoldChange))
kid_lfc <- kid_lfc %>% rename(ensembl = V1,padj_kid = V2,log2FoldChange_kid = V3)
sig_kid_lfc <- as.data.frame(subset(restrt_kid, padj < 0.05))
kid_lfc$ensembl <- as.factor(kid_lfc$ensembl)

hyp_lfc <- as.data.frame(cbind(rownames(restrt_hyp),restrt_hyp$padj, restrt_hyp$log2FoldChange))
hyp_lfc <- hyp_lfc %>% rename(ensembl = V1,padj_hyp = V2,log2FoldChange_hyp = V3)
sig_hyp_lfc <- as.data.frame(subset(restrt_hyp, padj < 0.05))
hyp_lfc$ensembl <- as.factor(hyp_lfc$ensembl)

gi_lfc <- as.data.frame(cbind(rownames(restrt_gi),restrt_gi$padj, restrt_gi$log2FoldChange))
gi_lfc <- gi_lfc %>% rename(ensembl = V1,padj_gi = V2,log2FoldChange_gi = V3)
sig_gi_lfc <- as.data.frame(subset(restrt_gi, padj < 0.05))
gi_lfc$ensembl <- as.factor(gi_lfc$ensembl)

liv_lfc <- as.data.frame(cbind(rownames(restrt_liv),restrt_liv$padj, restrt_liv$log2FoldChange))
liv_lfc <- liv_lfc %>% rename(ensembl = V1,padj_liv = V2,log2FoldChange_liv = V3)
sig_liv_lfc <- as.data.frame(subset(restrt_liv, padj < 0.05))
liv_lfc$ensembl <- as.factor(liv_lfc$ensembl)


lu_lfc <- as.data.frame(cbind(rownames(restrt_lu),restrt_lu$padj, restrt_lu$log2FoldChange))
lu_lfc <- lu_lfc %>% rename(ensembl = V1,padj_lu = V2,log2FoldChange_lu = V3)
sig_lu_lfc <- as.data.frame(subset(restrt_lu, padj < 0.05))
lu_lfc$ensembl <- as.factor(lu_lfc$ensembl)


annos$ensembl<-annos$gene

lfc_merge1 <- merge(kid_lfc, hyp_lfc, by = "ensembl",all= TRUE)
lfc_merge2 <- merge(lfc_merge1,gi_lfc,by = "ensembl",all = TRUE)
lfc_merge3 <- merge(lfc_merge2,liv_lfc,by = "ensembl",all = TRUE)
lfc_merge4 <- merge(lfc_merge3,lu_lfc,by = "ensembl",all= TRUE)
lfc_merge <- merge(lfc_merge4,annos,by = "ensembl",all = TRUE)
lfc_merge$ensid <- NULL
lfc_merge$ensembl <- NULL
lfc_merge$transcript <- NULL
lfc_merge$Gene <- NULL
#lfc_merge <- lfc_merge %>% distinct(gene, .keep_all = TRUE)
#rownames(lfc_merge) <- lfc_merge$gene


data_FC <- as.data.frame(cbind(lfc_merge$gene,lfc_merge$log2FoldChange_kid,lfc_merge$log2FoldChange_hyp, lfc_merge$log2FoldChange_gi, lfc_merge$log2FoldChange_liv, lfc_merge$log2FoldChange_lu))

data_FC <- data_FC %>% rename(gene = V1,kid = V2,hyp = V3,gi=V4,liv=V5, lu=V6)
data_file <- reshape2::melt(data_FC,measure.vars = c("kid", "hyp", "gi","liv","lu"))

data_Pval<- as.data.frame(cbind(lfc_merge$gene,lfc_merge$padj_kid,lfc_merge$padj_hyp, lfc_merge$padj_gi, lfc_merge$padj_liv, lfc_merge$padj_lu))

data_Pval <- data_Pval %>% rename(gene = V1,kid = V2,hyp = V3,gi=V4,liv=V5, lu=V6)
data_Pval <- reshape2::melt(data_Pval,measure.vars = c("kid", "hyp", "gi","liv","lu"))
data_file$pval <- as.numeric(data_Pval$value)
data_file$value <- as.numeric(data_file$value)

data_file= data_file[!apply(data_file[,2:3] == 0, 1, FUN = any, na.rm = TRUE),]
#rownames(data_file) <- NULL

data_file$labels=round(data_file$value,3)

theme<-theme(panel.background = element_blank(),panel.border=element_blank(),axis.line = element_line(colour = "black"),panel.grid.major = element_blank(),panel.grid.minor = element_blank(),
             strip.background=element_blank(),axis.text.x=element_text(colour="black",size=16),axis.text.y=element_text(colour="black",size=16),
             axis.ticks=element_line(colour="black"),plot.margin=unit(c(1,1,1,1),"line"))


grey= unique(subset(data_file,data_file$pval>=0.05)[c("variable", "labels","gene")])
color= unique(subset(data_file,data_file$pval<0.05)[c("variable", "labels","gene")])
grey$direction[grey$labels<0]="down"
grey$direction[grey$labels>=0]="up"

color$direction[color$labels<0]="down"
color$direction[color$labels>=0]="up"

cbPalette = c("darkgrey"="darkgrey","down"="#721f81","up"="#fd9668")

vol_plot1 <- ggplot(data = grey, aes(x = labels, y = variable)) +
  geom_jitter(data = grey , aes(color = "darkgrey"),size = 1.25,height = 0.3) +
  geom_jitter(data = color,aes(x = labels, y = variable, color=direction), size = 1.3,height = 0.3) +
  scale_color_manual(values = cbPalette, labels = c("nonsignificant",'down','up'),name = 'direction') +
  scale_x_continuous(breaks = pretty(data_file$value, n = 10)) +  labs(title= "", x= "\nLog2FC", y= "") + 
  #coord_cartesian(xlim = c(-7, 5)) +
  theme(
    axis.title.x = element_text( size=16 ),
    axis.title.y = element_text(size=16),
    legend.text = element_text(size=16),
    legend.title = element_text(size=16)) +
  ggtitle("standard vs lowfat")+
  #geom_text_repel(aes(label = gene), hjust=0, vjust=0,size=3,parse=TRUE)+
  theme 
vol_plot1
ggsave("vol_plot_diet.png", width = 10, height = 10, dpi = 600)
```

```{r}
summary(ressex_kid_lf)
summary(ressex_hyp_lf)
summary(ressex_gi_lf)
summary(ressex_liv_lf)
summary(ressex_lu_lf)


kid_lfc <- as.data.frame(cbind(rownames(ressex_kid_lf),ressex_kid_lf$padj, ressex_kid_lf$log2FoldChange))
kid_lfc <- kid_lfc %>% rename(ensembl = V1,padj_kid = V2,log2FoldChange_kid = V3)
kid_lfc$ensembl <- as.factor(kid_lfc$ensembl)

hyp_lfc <- as.data.frame(cbind(rownames(ressex_hyp_lf),ressex_hyp_lf$padj, ressex_hyp_lf$log2FoldChange))
hyp_lfc <- hyp_lfc %>% rename(ensembl = V1,padj_hyp = V2,log2FoldChange_hyp = V3)
hyp_lfc$ensembl <- as.factor(hyp_lfc$ensembl)

gi_lfc <- as.data.frame(ressex_gi_lf)
gi_lfc <- gi_lfc %>% rename(padj_gi = padj,log2FoldChange_gi = log2FoldChange)
gi_lfc <- tibble::rownames_to_column(gi_lfc, "ensembl")
gi_lfc$ensembl <- as.factor(gi_lfc$ensembl)

liv_lfc <- as.data.frame(cbind(rownames(ressex_liv_lf),ressex_liv_lf$padj, ressex_liv_lf$log2FoldChange))
liv_lfc <- liv_lfc %>% rename(ensembl = V1,padj_liv = V2,log2FoldChange_liv = V3)
liv_lfc$ensembl <- as.factor(liv_lfc$ensembl)

lu_lfc <- as.data.frame(cbind(rownames(ressex_lu_lf),ressex_lu_lf$padj, ressex_lu_lf$log2FoldChange))
lu_lfc <- lu_lfc %>% rename(ensembl = V1,padj_lu = V2,log2FoldChange_lu = V3)
lu_lfc$ensembl <- as.factor(lu_lfc$ensembl)

#annos$ensembl<-annos$ensid

lfc_merge1 <- merge(kid_lfc, hyp_lfc, by = "ensembl",all = TRUE)
lfc_merge2 <- merge(lfc_merge1,gi_lfc,by = "ensembl",all = TRUE)
lfc_merge3 <- merge(lfc_merge2,liv_lfc,by = "ensembl",all = TRUE)
lfc_merge4 <- merge(lfc_merge3,lu_lfc,by = "ensembl",all = TRUE)
lfc_merge <- merge(lfc_merge4,annos,by = "ensembl",all = TRUE)
lfc_merge$ensid <- NULL
lfc_merge$ensembl <- NULL
lfc_merge$transcript <- NULL
lfc_merge$Gene <- NULL
#lfc_merge <- lfc_merge %>% distinct(gene, .keep_all = TRUE)
#rownames(lfc_merge) <- lfc_merge$gene
#lfc_merge$gene <- NULL



data_FC <- as.data.frame(cbind(lfc_merge$gene,lfc_merge$log2FoldChange_kid,lfc_merge$log2FoldChange_hyp, lfc_merge$log2FoldChange_gi, lfc_merge$log2FoldChange_liv, lfc_merge$log2FoldChange_lu))

data_FC <- data_FC %>% rename(gene = V1,kid = V2,hyp = V3,gi=V4,liv=V5, lu=V6)
data_file <- reshape2::melt(data_FC,measure.vars = c("kid", "hyp", "gi","liv","lu"))

data_Pval<- as.data.frame(cbind(lfc_merge$gene,lfc_merge$padj_kid,lfc_merge$padj_hyp, lfc_merge$padj_gi, lfc_merge$padj_liv, lfc_merge$padj_lu))

data_Pval <- data_Pval %>% rename(gene = V1,kid = V2,hyp = V3,gi=V4,liv=V5, lu=V6)
data_Pval <- reshape2::melt(data_Pval,measure.vars = c("kid", "hyp", "gi","liv","lu"))
data_file$pval <- as.numeric(data_Pval$value)
data_file$value <- as.numeric(data_file$value)

data_file= data_file[!apply(data_file[,2:3] == 0, 1, FUN = any, na.rm = TRUE),]
#rownames(data_file) <- NULL

data_file$labels=round(data_file$value,3)

theme<-theme(panel.background = element_blank(),panel.border=element_blank(),axis.line = element_line(colour = "black"),panel.grid.major = element_blank(),panel.grid.minor = element_blank(),
             strip.background=element_blank(),axis.text.x=element_text(colour="black",size=16),axis.text.y=element_text(colour="black",size=16),
             axis.ticks=element_line(colour="black"),plot.margin=unit(c(1,1,1,1),"line"))


grey= unique(subset(data_file,data_file$pval>=0.05)[c("variable", "labels","gene")])
color= unique(subset(data_file,data_file$pval<0.05)[c("variable", "labels","gene")])
grey$direction[grey$labels<0]="down"
grey$direction[grey$labels>=0]="up"

color$direction[color$labels<0]="down"
color$direction[color$labels>=0]="up"

cbPalette = c("darkgrey"="darkgrey","down"="#721f81","up"="#fd9668")

vol_plot2 <- ggplot(data = grey, aes(x = labels, y = variable,name=gene)) +
  geom_jitter(data = grey , aes(color = "darkgrey"),size = 1.25,height = 0.3) +
  geom_jitter(data = color,aes(x = labels, y = variable, color=direction), size = 1.3,height = 0.3) +
  scale_color_manual(values = cbPalette, labels = c("nonsignificant",'down','up'),name = 'direction') +
  scale_x_continuous(breaks = pretty(data_file$value, n = 10)) +  labs(title= "", x= "\nLog2FC", y= "") + 
  #coord_cartesian(xlim = c(-7, 5)) +
  theme(
    axis.title.x = element_text( size=16 ),
    axis.title.y = element_text(size=16),
    legend.position="none") +
  ggtitle("male vs female, low fat diet")+
  #geom_text_repel(aes(label = gene), hjust=0, vjust=0,size=3,parse=TRUE)+
  theme 
vol_plot2

ggsave("vol_plot_sex_lf.png", width = 10, height = 10, dpi = 600)

```

res_sex_one_diet
```{r}
summary(ressex_kid_sd)
summary(ressex_hyp_sd)
summary(ressex_gi_sd)
summary(ressex_liv_sd)
summary(ressex_lu_sd)


kid_lfc <- as.data.frame(cbind(rownames(ressex_kid_sd),ressex_kid_sd$padj, ressex_lu_sd$log2FoldChange))
kid_lfc <- kid_lfc %>% rename(ensembl = V1,padj_kid = V2,log2FoldChange_kid = V3)
kid_lfc$ensembl <- as.factor(kid_lfc$ensembl)

hyp_lfc <- as.data.frame(cbind(rownames(ressex_hyp_sd),ressex_hyp_sd$padj, ressex_hyp_sd$log2FoldChange))
hyp_lfc <- hyp_lfc %>% rename(ensembl = V1,padj_hyp = V2,log2FoldChange_hyp = V3)
hyp_lfc$ensembl <- as.factor(hyp_lfc$ensembl)

gi_lfc <- as.data.frame(cbind(rownames(ressex_gi_sd),ressex_gi_sd$padj, ressex_gi_sd$log2FoldChange))
gi_lfc <- gi_lfc %>% rename(ensembl = V1,padj_gi = V2,log2FoldChange_gi = V3)
gi_lfc$ensembl <- as.factor(gi_lfc$ensembl)

liv_lfc <- as.data.frame(cbind(rownames(ressex_liv_sd),ressex_liv_sd$padj, ressex_liv_sd$log2FoldChange))
liv_lfc <- liv_lfc %>% rename(ensembl = V1,padj_liv = V2,log2FoldChange_liv = V3)
liv_lfc$ensembl <- as.factor(liv_lfc$ensembl)


lu_lfc <- as.data.frame(cbind(rownames(ressex_lu_sd),ressex_lu_sd$padj, ressex_lu_sd$log2FoldChange))
lu_lfc <- lu_lfc %>% rename(ensembl = V1,padj_lu = V2,log2FoldChange_lu = V3)
lu_lfc$ensembl <- as.factor(lu_lfc$ensembl)

#annos$ensembl<-annos$ensid

lfc_merge1 <- merge(kid_lfc, hyp_lfc, by = "ensembl",all = TRUE)
lfc_merge2 <- merge(lfc_merge1,gi_lfc,by = "ensembl",all = TRUE)
lfc_merge3 <- merge(lfc_merge2,liv_lfc,by = "ensembl",all = TRUE)
lfc_merge4 <- merge(lfc_merge3,lu_lfc,by = "ensembl",all = TRUE)
lfc_merge <- merge(lfc_merge4,annos,by = "ensembl",all = TRUE)
lfc_merge$ensid <- NULL
lfc_merge$ensembl <- NULL
lfc_merge$transcript <- NULL
lfc_merge$Gene <- NULL
#lfc_merge <- lfc_merge %>% distinct(gene, .keep_all = TRUE)
#rownames(lfc_merge) <- lfc_merge$gene


data_FC <- as.data.frame(cbind(lfc_merge$gene,lfc_merge$log2FoldChange_kid,lfc_merge$log2FoldChange_hyp, lfc_merge$log2FoldChange_gi, lfc_merge$log2FoldChange_liv, lfc_merge$log2FoldChange_lu))

data_FC <- data_FC %>% rename(gene = V1,kid = V2,hyp = V3,gi=V4,liv=V5, lu=V6)
data_file <- reshape2::melt(data_FC,measure.vars = c("kid", "hyp", "gi","liv","lu"))

data_Pval<- as.data.frame(cbind(lfc_merge$gene,lfc_merge$padj_kid,lfc_merge$padj_hyp, lfc_merge$padj_gi, lfc_merge$padj_liv, lfc_merge$padj_lu))

data_Pval <- data_Pval %>% rename(gene = V1,kid = V2,hyp = V3,gi=V4,liv=V5, lu=V6)
data_Pval <- reshape2::melt(data_Pval,measure.vars = c("kid", "hyp", "gi","liv","lu"))
data_file$pval <- as.numeric(data_Pval$value)
data_file$value <- as.numeric(data_file$value)

data_file= data_file[!apply(data_file[,2:3] == 0, 1, FUN = any, na.rm = TRUE),]
rownames(data_file) <- NULL

data_file$labels=round(data_file$value,3)

theme<-theme(panel.background = element_blank(),panel.border=element_blank(),axis.line = element_line(colour = "black"),panel.grid.major = element_blank(),panel.grid.minor = element_blank(),
             strip.background=element_blank(),axis.text.x=element_text(colour="black",size=16),axis.text.y=element_text(colour="black",size=16),
             axis.ticks=element_line(colour="black"),plot.margin=unit(c(1,1,1,1),"line"))


grey= unique(subset(data_file,data_file$pval>=0.05)[c("variable", "labels","gene")])
color= unique(subset(data_file,data_file$pval<0.05)[c("variable", "labels","gene")])
grey$direction[grey$labels<0]="down"
grey$direction[grey$labels>=0]="up"

color$direction[color$labels<0]="down"
color$direction[color$labels>=0]="up"

cbPalette = c("darkgrey"="darkgrey","down"="#721f81","up"="#fd9668")

vol_plot3 <- ggplot(data = grey, aes(x = labels, y = variable,name=gene)) +
  geom_jitter(data = grey , aes(color = "darkgrey"),size = 1.25,height = 0.3) +
  geom_jitter(data = color,aes(x = labels, y = variable, color=direction), size = 1.3,height = 0.3) +
  scale_color_manual(values = cbPalette, labels = c("nonsignificant",'down','up'),name = 'direction') +
  scale_x_continuous(breaks = pretty(data_file$value, n = 10)) +  labs(title= "", x= "\nLog2FC", y= "") + 
  #coord_cartesian(xlim = c(-7, 5)) +
  theme(
    axis.title.x = element_text( size=16 ),
    axis.title.y = element_text(size=16),
    legend.position="none") +
  ggtitle("male vs female, standard diet")+
  #geom_text_repel(aes(label = gene), hjust=0, vjust=0,size=3,parse=TRUE)+
  theme 
vol_plot3
ggsave("vol_plot_sex_sd.png", width = 15, height = 10, dpi = 600)

vol_plot3|vol_plot2|vol_plot1
ggsave("vol_plot_all.png", width = 20, height = 10, dpi = 400)

```

```{r}
CCA_outliers=rownames(out)
DE_lu=rownames(sig_lu_lfc)
DE_liv=rownames(sig_liv_lfc)
DE_gi=rownames(sig_gi_lfc)
DE_hyp=rownames(sig_hyp_lfc)
DE_kid=rownames(sig_kid_lfc)
lu_wgcna=unique(combo_lu$gene)
liv_wgcna=unique(combo_liv$gene)
gi_wgcna=unique(combo_gi$gene)
hyp_wgcna=unique(combo_hyp$gene)
kid_wgcna=unique(combo_kid$gene)


intersect_lu <- Reduce(intersect, list(CCA_outliers,DE_lu,lu_wgcna))
intersect_liv <- Reduce(intersect, list(CCA_outliers,DE_liv,liv_wgcna))
intersect_gi <- Reduce(intersect, list(CCA_outliers,DE_gi,gi_wgcna))
intersect_hyp <- Reduce(intersect, list(CCA_outliers,DE_hyp,hyp_wgcna))
intersect_kid <- Reduce(intersect, list(CCA_outliers,DE_kid,kid_wgcna))

library(AnnotationDbi)
library(org.Mm.eg.db)

# 1:1 mapping symbol→entrez
entrez_ids <- mapIds(org.Mm.eg.db, gene_symbols, 'ENTREZ', 'SYMBOL')

# 1:n mapping symbol→…
print(columns(org.Mm.eg.db))  # check what you can query
df <- select(org.Mm.eg.db, gene_symbols, c(...), 'SYMBOL')

#BiocManager::install("ComplexHeatmap")
lu <- list(CCA_outliers=rownames(out),
            DE_lu=rownames(sig_lu_lfc),
            lu_wgcna=unique(combo_lu$gene))

ComplexHeatmap::make_comb_mat(lu,mode = "intersect")

liv <- list(CCA_outliers=rownames(out),
            DE_liv=rownames(sig_liv_lfc),
            liv_wgcna=unique(combo_liv$gene))

ComplexHeatmap::make_comb_mat(liv,mode = "intersect")

kid <- list(CCA_outliers=rownames(out),
            DE_kid=rownames(sig_kid_lfc),
            kid_wgcna=unique(combo_kid$gene))

ComplexHeatmap::make_comb_mat(kid,mode = "intersect")

gi <- list(CCA_outliers=rownames(out),
            DE_gi=rownames(sig_gi_lfc),
            gi_wgcna=unique(combo_gi$gene))

ComplexHeatmap::make_comb_mat(gi,mode = "intersect")

hyp <- list(CCA_outliers=rownames(out),
            DE_hyp=rownames(sig_hyp_lfc),
            hyp_wgcna=unique(combo_hyp$gene))

ComplexHeatmap::make_comb_mat(hyp,mode = "intersect")

intersect_lu
intersect_liv
intersect_gi
intersect_hyp
intersect_kid

got_con_gi <-gost(query =intersect_gi, 
                organism = "hsapiens", ordered_query = FALSE, 
                multi_query = FALSE, significant = TRUE, exclude_iea = FALSE, 
                measure_underrepresentation = FALSE, evcodes = TRUE, 
                user_threshold = 0.05, correction_method = "g_SCS", 
                domain_scope = "annotated", custom_bg = NULL, 
                numeric_ns = "", sources = "GO", highlight = TRUE)

head(got_con_gi$result, 10)


go_con_gi_res <- as.data.frame(cbind(got_con_gi$result$source,
                                     got_con_gi$result$term_id,
                                     got_con_gi$result$term_name,
                                     got_con_gi$result$intersection,
                                     got_con_gi$result$intersection_size,
                                     got_con_gi$result$p_value
                                     ))



go_con_gi_res <- top_n(go_con_gi_res, 10, V6)

liver_circle_data <- read.csv("~/Documents/UNH/rnaseq/diet_final/liver_circle_data.csv")
lu_circle_data <- read.csv("~/Documents/UNH/rnaseq/diet_final/lu_circle_data.csv")
hyp_circle_data <- read.csv("~/Documents/UNH/rnaseq/diet_final/hyp_circle_data.csv")
kid_circle_data <- read.csv("~/Documents/UNH/rnaseq/diet_final/kid_circle_data.csv")

list2df <- separate_rows(kid_circle_data, "gene", convert = TRUE)
#names(list2df) <- c('category', 'ID', 'term', 'gene','count',"adj_pval")


#restrt_gi_df <- as.data.frame(restrt_gi)
#
#genelist<-restrt_gi_df %>% filter(
#  rownames(restrt_gi_df) %in% list2df$gene)
#
#genelist <- tibble::rownames_to_column(genelist, "gene")
#
#circ <- merge(list2df,genelist,by="gene") 
#
#circ$baseMean <- NULL
#circ$lfcSE <- NULL
#circ$padj <- NULL
#circ$pvalue <- NULL
#circ$category <- NULL
#circ$ID <- NULL
#circ$count <- NULL
#circ$adj_pval <- NULL
#circ$zscore <- NULL
#circ$stat <- NULL
#
#names(circ) <- c('gene','term',"logFC")
#col_order <- c("term", "gene", "logFC")
#circ <- circ[, col_order]
#
#circ$logFC <- as.numeric(circ$logFC)


bool_matrix <-table(list2df)

dim(bool_matrix)

library(circlize)
circos.clear()
circos.par(gap.after=c(rep(1,length(rownames(bool_matrix))-1),10,rep(1,length(colnames(bool_matrix))-1),10))
#chordDiagram(bool_matrix)

chordDiagram(bool_matrix, annotationTrack = "grid", 
         preAllocateTracks = list(track.height = 0.3))

circos.trackPlotRegion(track.index = 1, panel.fun = function(x, y) {
xlim = get.cell.meta.data("xlim")
ylim = get.cell.meta.data("ylim")
sector.name = get.cell.meta.data("sector.index")
circos.text(CELL_META$xcenter, 
            ylim[1] + cm_h(0.5),
            sector.name,
            facing = "clockwise",
            niceFacing = TRUE, 
            adj = c(0, 0.5))


#circos.axis(h = "bottom",labels.cex = .6,sector.index = sector.name)
}, bg.border = NA)

```


```{r}
immuneLu <- c("CNTN2", "NFASC", "CD1D", "ARG1", "CCL11")
immuneliv <- c("A1BG", "C8A", "C9", "CRP", "ITIH4", "PCDHGB1", "ULBP1")
immunehyp <- c("PTPRZ1", "RNF182", "ALPI", "EPCAM", "IGKC" )
immunekid <- c("PIGR")
immunegi <- c("CCL28", "IL7R")

inflammationLu <- c("ITIH1", "CCL11", "PYCARD", "SCGB1A1", "SCGB3A2" )
inflammationliv <- c("ITIH1")
inflammationgi <- c("CAMK2B", "ASIC4", "PRKCB", "ADCY2", "ADCY8")

circ <- c("GNAO1", "CAMK2B", "RYR2", "GNG3", "PRKCB", "ADCY2", "ADCY8", "CACNA1G", "GRIN1")

angiogensis <- c("ANG","ADTRP","ADTRP")
sodium <- c("SLC9A5","TNR",	"SLC38A4", "SLC17A1")
apoptosis<- c("AATK", "FAIM2", "EMP1", "PYCARD",	"APCS",		"XKR9")
glucose <- c("ITLN1",	"SLC2A9",		"GJB1","ADCYAP1R1","PKLR",	"PPP1R3G")
raas <- c("AGT", "ACE2", "CMA1", "AGTR1" )
vasso <- c("STX4", "RAB5C", "RAB11B", "CREB3L2", "AQP3", "VAMP2", "DYNLL2", "GNAS", "AVPI1")
#"AQP2"
gluconeogenesis <- c("CYP2E1", "PFKFB1", "PKLR","PPP1R3G", "ENO2","ADH4","ADH6","ALDOB","FBP1","PCK1","PKLR","HKDC1")
insulin <- c("PKLR", "PPP1R3B", "HKDC1", "PCK1", "FBP1", "RIMS2", "CAMK2B", "RYR2", "SNAP25", "CREB3L3", "PRKCB", "KCNMA1", "SLC2A2", "ADCY2", "ATP1A2", "GCG", "ADCY8")
gluconeogenesisGI <- c("ADH4", "PKLR", "HKDC1", "ALDOB", "ENO2", "PCK1", "FBP1", "ADH6")
insulinL <- c("HNF4A", "IGFALS", "ITIH2")
coag <- c("C8G", "CFI", "PLG", "F13B", "F5")
dig_abs_fat_GI <- c("FABP1", "PNLIPRP2", "ABCG8", "FABP2", "ABCG5", 'PLA2G12B', "NPC1L1", "MTTP", "APOA4", "APOB","PDK1","CPT1A")
dig_abs_carb_GI <- c("PRKCB", "HKDC1", "SLC2A2", "ATP1A2","PDHA1","SLC5A1")
pdk_pdh <- c("PDK1", "PDHA1")

datalist = list()

for (i in dig_abs_carb_GI) 
  {
counts <- plotCounts(ddstrt, gene = i, intgroup = c("trt", "sex", "tissue"), returnData = TRUE, xlab="trt")
counts$gene <- i
datalist[[i]] <- rbind(counts)
}

big_data = do.call(rbind, datalist)

big_data$trt <- as.factor(big_data$trt)
big_data$sex <- as.factor(big_data$sex)
big_data$gene <- as.factor(big_data$gene)
big_data$tissue <- as.factor(big_data$tissue)
big_data$count <- as.numeric(big_data$count)


cols <- c("kid"="#cde11d","liv"="#4cc26c","lu"="#20938c", "hyp"="#6c00a8", "gi" = "#3d4d8a")

coo_genes <- ggplot(big_data,  aes(x = trt, y = log10(count), fill = tissue, alpha = trt)) + 
      #ggtitle(title)  + 
      theme +
      ylab("Normalized counts (Log10 transformed)") + 
      xlab("diet") +
      geom_violin(width=1.4) +
      geom_boxplot(width=0.1, color="grey", alpha=0.2) +
      facet_grid(gene~tissue , scales = "free") +
      theme_classic() +
      theme(panel.border = element_rect(color = "black", fill = NA, size = .5),  legend.title=element_text(size=10), legend.text=element_text(size=10),axis.text.x=element_text(colour="black",size=10),axis.text.y=element_text(colour="black",size=10))+
      guides(alpha = FALSE)+
      scale_fill_manual(values=cols, name = "tissue")+
      scale_alpha_discrete(range=c(0.6, 1))
coo_genes
```

